CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-08 codecheck-paper.tex
done

No approach besides CODECHECK spans across journals and is
intentionally designed to be adopted across journals or events and to
build a community of codecheckers. CODECHECK shares its
interdisciplinary nature with other community initiatives concerned
with reproducibility awareness, education, and support, such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaboration with the preprint server
\emph{arXiv} on providing data and code supplements for machine
learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
and runs a reproduciblity
challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
GIScience \cite{nust_agile_2019}, which naturally share some aspects
while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for
authors, we see CODECHECK's role as referencing and linking
educational efforts and helpful material, not as creating and
maintaining such content.

@@

\textbf{Isn't CODECHECK what peer review should be doing already?} \\
On the surface, yes.
However, the general observation is that the
increasing number of researchers, submissions, and reviews to conduct
under the pressure to publish puts the scholarly publication system
under significant strain.
Miyakawa described how when he as an editor simply requested raw data from authors, this led to almost all of those papers in which data was requested to be withdrawn or rejected \cite{miyakawa_no_2020}.
Such basic checks to show that
sufficient data supports the results requires much effort from editors,
yet they only rely on the availability of data files and the content of the
paper.
These availability checks can be enhanced by having more complex CODECHECKs request the code and then execute it, because code is not inert but needs to be run in a specific environment.
This might fall within idealistic expectations of peer review,
but is far from common in reality.
Establishing a CODECHECK
process is largely an acknowledgement that peer reviewing practices
have been unable to adapt to the challenges of digitisation of data-based science.
The concept of a CODECHECK,
just as the concepts of reproducible research and Open Science,
may be transitional by nature. If the activities described here as
being part of a CODECHECK become the norm for scientific reviews and
publications, then the initiative will have succeeded. However, the
rapidly changing technology used in research leads us to believe that
a separate role of codechecker and the expertise it brings will be
beneficial in the long term.

@@

\textbf{Should CODECHECK requirements not be more demanding?} \\
CODECHECK by design does not require authors to provide (and sustain) an
eternally functional workflow and neither suggests a specific software stack
or practical approach even if such suggestions, e.g.,~research compendia or ``repro-packs''\footnote{See \url{https://research-compendium.science/} respectively \cite{barba_praxis_2018}.}, are very valuable.
Creating something that anyone can reproduce has been called a 
fool's errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.
However, the package of data, code, and documentation
collaboratively created by authors and codecheckers is a snapshot of a 
working analysis that greatly increases the likelihood of a successful 
reproduction and the possibility that a workflow can be extended by third parties in the future, as long as they have access
to suitable resources and matching skill set.
Concrete implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors on how to
create reproducibility packages.
The author-friendly ``low bar'' should not stay low forever, but cultural
change takes time and the encouragement and guidance that CODECHECK,
as part of the widely accepted peer review concept, can provide may
eventually allow the bar to be raised as high as possible, e.g.,~with
executable research compendia \cite{nust_opening_2017}, ``Whole Tales'' \cite{brinckman_computing_2018},
or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
However, considering that missing artefacts and lack of documentation
have repeatedly been identified as key barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
we would not underestimate the power of a simple check.
For example, ModelDB curation policies ensure that only one figure is manually reproduced \cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness nor success of the platform.

It should be noted that a codechecker does not fulfil the same role as a
\emph{statistical reviewer}, as it is applied by some journals in the 
biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
The statistical reviewer actually evaluates the appropriateness of
statistical methods \cite{greenwood_how_2015} and can support topical
reviewers if, e.g.,~complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.
The codechecker may go equally deep into the review, if they have
the required expertise and the required time, 
but they are by no means expected to be able to provide
such a level of scrutiny. We can imagine a tiered workflow where a 
codechecker could, just as a conventional reviewer could, recommend a detailed
statistical review to the editor if they come upon certain issues while
examining the work from a different perspective.

An important distinction is that a codechecker does not conduct a 
\emph{code review}. Code reviews are without a doubt extremely valuable in 
terms of improving reproducibility and reusability, and their proponents 
even believe they can improve the science \cite{petre_code_2014}.
Code reviews, however, have quite different structural challenges and 
require even more resources. That said, a well-reviewed codebase is likely
to be easier to codecheck, and the awareness of high-quality code raised
through CODECHECK may lead to more support for code reviewing.
Initiatives and journals that conduct software reviews independent of 
a specific publication or workflow include
ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
and JOSS\footnote{\url{https://joss.theoj.org/}}.
Furthermore, the codechecker's task list is intentionally not
overloaded with related concerns, albeit potentially important ones,
such as ensuring proper citation of data and software or depositing
data and software in suitable repositories. Nevertheless,
we expect codecheckers to often recognise these issues and highlight them.

\textbf{How are failures during checks handled?}
We have no established process for the case
in which a reproduction completey fails.
This may be a sign that our pre-selection of the examples was biased, since so far all our CODECHECKs succeeded.
In the case that a journal adopts CODECHECK for all submissions, the open 
question remains as what to do if a check fails.
A negative comment in a CODECHECK certificate or a failed check does
not necessarily mean the paper or the science is bad
(cf. discussion on negative comments in \cite{everythinghertz123}).
At the start, we doubt that publicly reporting failures (i.e.,~the code 
would not run) will increase overall reproducibility, and we recommend not
sharing that outcome publicly but sharing the latest draft of the 
certificate with the authors.
One must be careful how policies on disclosing failures may negatively impact the acceptance by authors, but also how efforts by volunteers may go unnoticed if no report is published.
The most appropriate way forwards seems to be to give private feedback for failures and public acknowledgement and encouragement for successes.
% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one way for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

\textbf{Who will pay for the compute time?}
For those papers that take significant compute
time (days, not minutes), it is unclear who will pay for it.
The openness of CODECHECK certificates could help to establish community
peer pressure, or recognition, as submitting authors can be evaluated on their engagement
as codecheckers. This may lead to a balance of the shared load to reproduce
computations.
However, one must carfully consider the sustainability of rerunning computations and the
environmental impact large calculations, such as training machine learning models, have.
The challenges of computational resources can also quickly be reduced when
authors provide a small sample or a synthetic dataset together with the
expected outputs for this small example, which can equally well demonstrate the 
usefulness and validity of their workflow.

\textbf{What about my proprietary software and sensitive data?}
Authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, e.g MATLAB, this seems like a pragmatic choice.
However, this prohibits publications from benefiting from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
and requires the codechecker to have access to that particular software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

Solutions for proprietary and sensitive data exists.
Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
independent entities can provide infrastructure for sharing data and workflows
confidentially \cite{perignon_certify_2019} or with access to derived results
but not raw data \cite{shannon_opening_2018},
i.e.,~data enclaves \cite{foster_research_2018},
or domains of reproducibility \cite{harris_more_2017})

\textbf{Can't someone cheat?} Yes. We don't check that the code is correct or
sound science,
just simply that it runs. This ``mechanical'' test is indeed a very low bar.
Yet, by having the code and (raw) data openly deposited, interested parties can
then examine the code, and we hope that knowing the code will be open will
be an incentive for authors to ensure good quality.
It will also give opportunities
to future researchers, potentially with new methods, to look for errors.
This prospective evaluation is a much more effective way to capture cheaters than engaging in an arms race on tools, methods, or resources trying to catch malicious intent now with closed datasets and code; this approach is comparable to storing blood samples of sport champions today to possibly detect doping in the future with more advanced methods (cf. \cite{everythinghertz97}).
Another comparison that helped us define the scope of a CODECHECK is was that we think of the codechecker as forensic photographer, making sure all the details
are captured so that an investigator may, now or in the future, scrutinize
them.

\textbf{Who's got time for more peer review?} Agree.
The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
However, a CODECHECK is different from peer review.
First, the \textbf{technical nature} of a CODECHECK can provide a more 
clearly defined scope and expectations and thereby time budget compared to conventional 
peer review. 
For example, authors are told what to provide and the codechecker can be told when to stop.
The option for the codechecker to ask for clarification or more 
material can greatly reduce the digging required and thereby
increase \textbf{efficiency}.
Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
Research software experts (Research Software Engineers, RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}}) who might
not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
In fact, the codechecker could even be a software professional employed by a publisher.
Third, recall that we expect CODECHECKs not to be anonymous and always
publicly deposited, which is common practice for reviews of few journals only.
With code and workflows, there also is a good chance that a codechecker's
feedback can directly impact and improve the author's 
work, which may make for a rewarding experience or even provide tangible contributions.
The public certificates and contributions provide \textbf{peer recognition}.
Fourth, we also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very \textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the community.

@@

\textbf{Do workflows need to be codechecked multiple times?}
If you check a preprint before peer
review or a submission to a journal with non-public review, it might
need checking again as it is modified during peer review.
This is somewhat unavoidable, and has happened to us (e.g.,~certificate \texttt{2020-012}
\cite{cert-2020-012} has been revised). This can also be desired tough, if
the interaction between author, reviewer, and codechecker led to
improvements. Since checking the manuscript the second time is likely
to be much less work than the first time, this is largely mitigated by
open and regular communication. The higher the reproducibility and
automation, the more aspects of a CODECHECK that may eventually be
supported by automation, though we believe the human codecheckers
should remain the decision makers. Automatically detecting whether results
are ``the same'' is a question intrinsically linked with each unique
process and quite hard to realise for graphical outputs, such as plots or
maps.
That is why we do not require results to be exactly the same for a
CODECHECK to pass but simply that the code runs and generates the output
files that the author claims. Stochastic simulations mean that often
we cannot get exactly the same results, and that even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.
While reproducibility practices can mitigate some of these problems, e.g., by using a seed, the flexibility of the human judgement is still needed.

\textbf{Shouldn't the next step be more revolutionary?}
CODECHECK's approach is to acknowledge the shortcomings around 
computational reproducibility and to iteratively improve the current
system of scholarly communication based on journals and conferences.
It remains to be proven whether this approach is welcomed broadly and whether 
involving the traditional publishing stakeholders helps to further the cause.
We have discussed stronger more drastic rules at length, such as only 
considering fully free and open source software, diamond Open Access journals or community-led no-fee journals,
and journals with collaborative/open peer review, but we eventually decided
against them on the level of the principles.
For the CODECHECK community process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
and the volunteer codechecker community, these requirements are reconsidered.

We also have deliberated requiring the latest technologial solutions which can
support reproducibility (cf.~\cite{konkol_publishing_2020}), but we have 
decided to focus on the human interface and the judgement of experienced
researchers and developers as a more sustainable and powerful approach.
The structurelessness of the required data and code gives all types of research the freedom to adopt CODECHECK.
CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
Note that the focus of the CODECHECK principles on code execution allows journals
and publishers to innovate on the financial model and peer review practices
at their own pace.

@@

CODECHECK works!
We have created a considerable number of certificates to demonstrate it.
The creation of the certificates and the interactions with authors and
editors not only improved the principles and the workflow but also
confirmed the approach taken. This result corroborates findings from
similar evaluations of reproducible computational research in journals and
at conferences.
CODECHECKs increase transparency of the checked papers and can contribute
to building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of 
shared principles and common name, through recognition value,
will allow researchers
across publication venues to judge the level of scrutiny that results have
faced. CODECHECK requires direct acknowledgement of the contribution to the
scientific community in the role of codechecker, not indirectly via citations 
of reproductions or informal crediation and closed inner circles.
Forther, codechecking can be a worthwhile educational activity, especially for
ECRs.

Yet, CODECHECK harbours the same limitations as peer review
in general and is closely connected to larger disruptions and challenges
in scholarly communication
\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
including the tensions between commercial publishing and reviewers' often
free labour, and a global pandemic that has jumbled up scientific publishing
and exposed a broader general audience to science, even preliminary 
results \cite{munafo_what_2020}.
It is clear that establishing CODECHECK-like systems can be 
impeded by and must be seen as interconnected with much larger issues in 
science, such as broken metrics or malpractice triggered by publication
pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
We certainly do not want the binary attribute of ``code works''
to become a factor in bibliometric approaches for performance assesments.
% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745

While developed for the current ``paper''-centric publication process and
being compatible with different styles of peer review, the CODECHECK 
principles would also work well with novel publication paradigms, e.g.,
peer-reviewed computational notebooks \cite{earthcube_new_2020},
iterative and granular communication of scientific outputs, 
articles with live-code \cite{perkel_pioneering_2019-1}
\footnote{For example, eLife's ERA, see 
\url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019},
and completely new visions for scholarly communication and peer 
review\footnote{Cf., for example, \emph{A modern vision for peer review} by 
Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
An explicit segmentation of research steps could even make the focus 
of a CODECHECK easier by only checking the ``analysis'' sub-publication.
The discovery of CODECHECKs could be increased by depositing the reports
into public databases of reproductions, such as \emph{SciGen.Report}.
Public researcher profiles, such as ORCID, may consider different
types of reviewer activity to capture how independent code execution contributes to science.
Notably, the discussed limitations are largely self-imposed for easier
acceptance and evolutionary integration, as to not break the current 
system and increase demands gradually without leaving practitioners behind.

If adopted today, a CODECHECK system, even if temporary in a sustainable 
transition towards more open publication and review practices, can contribute
to increased trust in the results of research. Introducing CODECHECK
should be informed by lessons learned from (introducing) open peer review
(OPR) \cite{ross-hellauer_guidelines_2019}.
All our conversations with publishers and editors at some point
touched on the procedural and technical implications of openness, be it 
that OPR is neither applied by nor familiar to stakeholders, or that process
management software would not support such open communication.

More reproducible practices initiated by CODECHECKs could lead
communities or whole disciplines to reach a state where authors provide 
sufficient material and reviewers have acquired sufficient skills that
the reviewers will generally conduct a CODECHECK-level of checking;
only in especially sophisticated cases will a specialised codechecker
be needed.
The main challenge for us remains getting journals to embrace the
idea behind CODECHECK and to realise processes that conform to the principles,
whether or not they use CODECHECK by name.
We would be keen to use the flexibility of the principles and cooperate
with journals to learn more about the advantages and yet unclear
specific challenges -- Do CODECHECKs really work better with OPR?
To facilitate the adoption, the CODECHECK badge is, intentionally, not branded beyond the checkmark
and green colour and simply states ``code works''.

Each rollout of CODECHECK should be accompanied by studies to ensure 
codechecking does not fall into the same gaps as peer review did
\cite{tennant_limitations_2020}
and to ensure positive change within the review system.
This \emph{cultural change}, however, is needed for the valuation of the
efforts that go into proper evaluation of papers.
Multiple adopting journals will also help to answer relevant questions on
the establishing process: What are crucial decisions or pain points? Can
authors retract code/data once a CODECHECK has started?
What variants of CODECHECKs will be most common? How will open CODECHECKs
influence or codevelop with the scope and anonymity of conventional review over time.

>>>>

CODECHECK is uniquely designed to be adopted across journals or events
and to build a community of codecheckers. CODECHECK shares its
interdisciplinary nature with other community initiatives concerned
with reproducibility awareness, education, and support, such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaboration with the preprint server
\emph{arXiv} on providing data and code supplements for machine
learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
and runs a reproduciblity
challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
GIScience \cite{nust_agile_2019}, which naturally share some aspects
while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for
authors, we see CODECHECK's role as referencing and linking
educational efforts and helpful material, not as creating and
maintaining such content.

@@

\textbf{Isn't CODECHECK what peer review should be doing already?} \\
On the surface, yes, but peer reviewers are overburdened enough and
asking them to do more work usring peer review is not likely to
suceed.  Miyakawa described how when he as an editor simply requested
raw data from authors during submission, this led to almost all of
those papers in which data was requested to be withdrawn or rejected
after peer review \cite{miyakawa_no_2020}.  Such basic checks to show
that sufficient data supports the results requires effort from
editors, yet they only rely on the availability of data files and the
content of the paper.  These availability checks can be enhanced by
having more complex CODECHECKs request the code and then execute it.
This might fall within idealistic expectations of peer review, but is
rare.  Establishing a CODECHECK process is largely an acknowledgement
that peer reviewing practices have been unable to adapt to the
challenges of computational papers.  The concept of a CODECHECK, just
as the concepts of reproducible research and Open Science, may be
transitional by nature. If the activities described here as being part
of a CODECHECK become the norm for scientific reviews and
publications, then the initiative will have succeeded. However, the
rapidly changing technology used in research leads us to believe that
a separate role of codechecker and the expertise it brings will be
beneficial in the long term.

@@

\textbf{Should CODECHECK requirements be more demanding?} \\
CODECHECK by design does not require authors to provide (and sustain)
an eternally functional workflow nor suggests a specific software
stack or practical approach even if such suggestions, e.g.,~research
compendia or ``repro-packs''\footnote{See
  \url{https://research-compendium.science/} respectively
  \cite{barba_praxis_2018}.}, are valuable.  Creating something
that anyone can reproduce has been called a fool's
errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.  However, the package of data, code, and
documentation collaboratively created by authors and codecheckers is a
snapshot of a working analysis that greatly increases the likelihood
of a successful reproduction and the possibility that a workflow can
be extended by third parties in the future, as long as they have
access to suitable resources and matching skill set.  Concrete
implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors on how
to create reproducibility packages.  Our author-friendly ``low bar''
should not stay low forever, but cultural change takes time and the
encouragement and guidance that CODECHECK, as part of the widely
accepted peer review concept, can provide may eventually allow the bar
to be raised much higher, e.g.,~with executable research
compendia \cite{nust_opening_2017}, ``Whole Tales''
\cite{brinckman_computing_2018}, or continuous analysis
\cite{beaulieu-jones_reproducibility_2017-1}.  However, considering
that missing artefacts and lack of documentation have repeatedly been
identified as key barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}), we would not
underestimate the power of a simple check.  For example, ModelDB
curation policies require that only one figure need be manually reproduced
\cite{mcdougal_reproducibility_2016}, but that has not limited the
usefulness nor success of the platform.

A codechecker does not fulfil the same role as a \emph{statistical
  reviewer}, as it is applied by some journals in the biomedical
domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).  The
statistical reviewer evaluates the appropriateness of statistical
methods \cite{greenwood_how_2015} and can support topical reviewers
if, e.g.,~complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.  The codechecker may go
equally deep into the review, but only if they have the expertise and
time. We can imagine a tiered workflow where a codechecker could, just
as a conventional reviewer could, recommend a detailed code review
(see next paragraph) to the editor if they come upon certain issues while
examining the work.

A codechecker does not conduct a \emph{code review}. Code reviews are
valuable to improve reproducibility and reusability, and their
proponents even believe they can improve the science
\cite{petre_code_2014}.  Code reviews, however, have quite different
structural challenges and require even more resources. That said, a
well-reviewed codebase is likely to be easier to codecheck, and the
awareness of high-quality code raised through CODECHECK may lead to
more support for code reviewing.  Initiatives and journals that
conduct software reviews independent of a specific publication or
workflow include ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}}, and
JOSS\footnote{\url{https://joss.theoj.org/}}.  Furthermore, the
codechecker's task list is intentionally not overloaded with related
concerns, albeit potentially important ones, such as ensuring proper
citation of data and software or depositing data and software in
suitable repositories. Nevertheless, we expect codecheckers to often
recognise these issues and highlight them.

\textbf{How are failures during checks handled?}

We have not yet established a process for if a reproduction fails.  This
may simply reflect our biases in pre-selection of case studies.  In
the case that a journal adopts CODECHECK for all submissions, the open
question remains as what to do if a check fails, after exhausting
efforts between author and codechecker to reproduce the workflow.  A
negative comment in a CODECHECK certificate or a failed check does not
necessarily mean the paper or the science is bad (cf. discussion on
negative comments in \cite{everythinghertz123}).  We doubt that
publicly reporting failures (i.e.,~the code would not run) will
increase overall reproducibility, and we recommend sharing the latest
draft of the certificate only with the authors.  One must be careful
how policies on disclosing failures may \textbf{negatively impact the
  acceptance by authors}
%% SJE: what does this mean?
, and also how volunteer efforts may go unnoticed if no report is
published.  The most appropriate way forwards seems to be to give
private feedback for failures and public acknowledgement and
encouragement for successes.
% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one way for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

\textbf{Who will pay for the compute time?}  For papers that take
significant compute time (days, not minutes), it is unclear who will
pay for it. One must carefully consider the sustainability of
rerunning computations and the environmental impact large
calculations, such as training machine learning models, have.  A
pragmatic workaround is to request that authors provide a ``toy''
example, or small dataset that can be quickly analysed to demonstrate
that the workflow runs correctly.

\textbf{What about my proprietary software and sensitive data?}  Given
the prevalence of proprietary software, e.g MATLAB, we pragmatically
decided that we should accept code as long as we could find a machine
with suitable licences to run it.  However, this prohibits us from
using open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020,perkel_make_2019}) and requires the
codechecker to have access to that particular software.  Non-open
software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a
compromise that should be reconsidered.

Solutions for proprietary and sensitive data exists.  Authors can
provide synthetic data (cf. \cite{shannon_opening_2018}), some data
can effectively be redacted \cite{oloughlin_data_2015}, and publishers
or independent entities can provide infrastructure for sharing data
and workflows confidentially \cite{perignon_certify_2019} or with
access to derived results but not raw data
\cite{shannon_opening_2018}, i.e.,~data enclaves
\cite{foster_research_2018}, or domains of reproducibility
\cite{harris_more_2017}).

\textbf{Can't someone cheat?} Yes. We simply check that the code runs,
not that is correct or sound science. This ``mechanical'' test is
indeed a low bar.  By having the code and data openly deposited, third
parties can later examine the code, and we hope that knowing the code
will be open ensures that authors will not cheat.  It also allows
researchers, potentially with new methods, to look for errors.  This
is more effective than engaging in an arms race on building methods to
detect malicious intent now with closed datasets and code.  This is
analogous to storing blood samples of sport champions today to
possibly detect doping in the future with more sensitive methods
(cf. \cite{everythinghertz97}).  Another comparison that helped us
define the scope of a CODECHECK is that we think of the codechecker as
forensic photographer, capturing details so that an investigator may
later scrutinize them.

\textbf{Who's got time for more peer review?} Agree; codechecking
takes time that could otherwise be used for traditional peer review.
However, a CODECHECK is different from peer review.  First, the
\textbf{technical nature} of a CODECHECK sets clear expectations and
thereby time budget compared to conventional peer review.  For
example, authors are told what to provide and the codechecker can be
told when to stop.  Codecheckers can always directly ask the author
when clarification is required, thereby increasing
\textbf{efficiency}.  Second, the specific skill set of a codechecker
allows for \textbf{different groups} to participate in the review
process.  ECRs might be attracted to learn more about recent methods,
peer review, and reproducibility practices.  Research software experts
(Research Software Engineers,
RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}})
who might not regularly be involved in writing or reviewing papers
might be interested in increasing their connection with scholarly
practices.  An extra codechecker may simplify the matchmaking an
editor does when identifying suitable reviewers for a submission,
as technical and topical expertise can be provided by different
people.  Third, recall that CODECHECKs should always be publicly
available, unlike peer review reports.  With code and workflows, the
codechecker's feedback may directly impact and improve the author's
work.  The public certificates and contributions provide \textbf{peer
  recognition} for the codechecker.  Fourth, we found that focusing on
workflow mechanics and interacting with the author makes reproductions
\textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the
community.

@@

\textbf{Do workflows need to be codechecked multiple times?}  If a
paper is checked at the start of peer review, it might need
re-checking if the paper is modified during peer review.  This is
somewhat unavoidable, and has happened to us (e.g.,~certificate
\texttt{2020-012} \cite{cert-2020-012}). This can
also be desired though, if interactions between author, reviewer,
and codechecker led to improvements.  Checking the manuscript the
second time is likely to be much less work than the first time.

\textbf{What does it mean for a figure to be reproducible?}

Automatically detecting if a codechecker's results are ``the same'' as
an author's is more challenging than it might appear.  That is why we
do not require results to be \textit{identical} for a CODECHECK to pass
but simply that the code runs and generates output files that the
author claims. Stochastic simulations mean that often we cannot get
exactly the same results, and that even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.  While reproducibility practices can
mitigate some of these problems, e.g., by using a seed, the
flexibility of the human judgement is still needed, rather than
bitwise reproducibility.  The codechecker is free to comment on
visible differences in outputs in their report.

\textbf{Shouldn't the next step be more revolutionary?}  CODECHECK's
approach is to acknowledge shortcomings around computational
reproducibility and to iteratively improve the current system.  It
remains to be proven whether this approach is welcomed broadly and if
involving publishing stakeholders helps to further the cause.  We have
discussed more stringent rules at length, e.g. only considering fully
free and open source software, diamond Open Access journals, but we
eventually decided against them on the level of the principles.  For
the CODECHECK community
process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
and the volunteer codechecker community, these requirements are
reconsidered.

We also have deliberated requiring modern technological solutions to
support reproducibility (cf.~\cite{konkol_publishing_2020}), but we
decided to focus on the human interface and the judgement of
experienced researchers and developers as a more sustainable and
flexible approach.  All types of research can adopt CODECHECK due to
its flexible design.  CODECHECK could include automated scoring
(e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new
risks.  The focus of the CODECHECK principles on code execution allows
journals and publishers to innovate on financial  models and peer
review practices at their own pace.

@@

CODECHECK works --- we have created a considerable number of
certificates to demonstrate it.  The creation of certificates and
interactions with authors and editors shaped the principles and the
workflow and also confirmed the approach taken. This result
corroborates findings from similar evaluations of reproducible
computational research in journals and conferences.  CODECHECKs
increase transparency of the checked papers and can contribute to
building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of shared principles and common name, through recognition
value, will allow researchers to judge the level of scrutiny that
results have faced. CODECHECK requires direct acknowledgement of the
codechecker's contributions, not indirectly via citations of
reproductions or informal crediation and closed inner circles.

CODECHECK however harbours the same limitations as peer review in
general and is closely connected to larger disruptions and challenges
in scholarly communication
\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, including
the tensions between commercial publishing and reviewers' often free
labour, and a global pandemic that has jumbled up scientific
publishing and exposed a broader general audience to science, even
preliminary results \cite{munafo_what_2020}.  Establishing CODECHECK
workflows must be seen as interconnected with much larger issues in
science, such as broken metrics or malpractice triggered by
publication pressure
\cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.  We certainly do
not want the binary attribute of ``code works'' to become a factor in
bibliometric approaches for performance assesments.
% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745

While developed for the current ``paper''-centric publication process,
the CODECHECK principles would also work well with novel publication
paradigms, e.g., peer-reviewed computational notebooks
\cite{earthcube_new_2020}, iterative and granular communication of
scientific outputs, articles with live-code
\cite{perkel_pioneering_2019-1} \footnote{For example, eLife's ERA,
  see
  \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019}, and completely new visions for
scholarly communication and peer review\footnote{Cf., for example,
  \emph{A modern vision for peer review} by Amy J. Ko:
  \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
An explicit segmentation of research steps could even make the focus
of a CODECHECK easier by only checking the ``analysis''
sub-publication.  The discovery of CODECHECKs could be increased by
depositing the reports into public databases of reproductions, such as
\emph{SciGen.Report}.  Public researcher profiles, such as ORCID, may
consider different types of reviewer activity to capture how
independent code execution contributes to science.  Notably, the
discussed limitations are largely self-imposed for easier acceptance
and evolutionary integration, as to not break the current system and
increase demands gradually without leaving practitioners behind.

A CODECHECK system, even if temporarily adopted as a sustainable
transition towards more open publication and review practices, can
contribute to increased trust in research outputs. Introducing
CODECHECK should be informed by lessons learned from (introducing)
open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  Our
conversations with publishers and editors included procedural and
technical implications of openness, be it that OPR is neither applied
by nor familiar to stakeholders, or that process management software
would not support such open communication.

More reproducible practices initiated by CODECHECKs could lead
communities to reach a state where authors provide sufficient material
and reviewers have acquired sufficient skills that peer reviewers will
generally conduct a CODECHECK-level of checking; only in especially
sophisticated cases will a specialised codechecker be needed.  The
main challenge for us remains getting journals to embrace the idea
behind CODECHECK and to realise processes that conform to the
principles, whether or not they use CODECHECK by name.  We would be
keen to use the flexibility of the principles and cooperate with
journals to learn more about the advantages and yet unclear specific
challenges -- e.g do CODECHECKs really work better with OPR?  To
facilitate the adoption, the CODECHECK badge is, intentionally, not
branded beyond the checkmark and green colour and simply states ``code
works''.

Future CODECHECK versions may be accompanied by studies to ensure
codechecking does not fall into the same traps as peer review did
\cite{tennant_limitations_2020} and to ensure positive change within
the review system.  This \emph{cultural change}, however, is needed
for the valuation of the efforts that go into proper evaluation of
papers.  Journals can help us to answer open questions in our system: What are
crucial decisions or pain points? Can authors retract code/data once a
CODECHECK has started?  What variants of CODECHECKs will be most
common? How will open CODECHECKs influence or codevelop with the scope
and anonymity of conventional review over time?

The question of training codecheckers is also relevant. We expect a
mentoring scheme within the CODECHECK community (experienced
codecheckers will provide on-the-job training or serve as fallback
advisors). Codecheckers may also be found by collaborating with
reproducible research initiatives such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
\cite{fitzgibbon_brewing_2020}, and
Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
\cite{auer_reproducibility_2020}.  The initial reaction of the
scientific community to these ideas shows that scholarly peer review
should continue on the path towards facilitating the sharing and
execution of computational workflows.
